### 2014-12-08 Toy Example/expected value/mu_parameterization.Rmd
initial commit

---
author: "Andee Kaplan"
date: "11/08/2014"
output:
  html_document:
    theme: readable
runtime: shiny
---

# $\mu$-Parametrization
In Random Graph Model theory, degeneracy occurs when the model places a disproportionate probability on only a few elements of the sample space, $\mathcal{X}$. Handcock, et al. define a model to be near degenerate if $\mu(\boldsymbol \theta)$, the mean parametrization on the model parameters, is close to the boundary of the convex hull of $\{t({\bf x}): {\bf x} \in \mathcal{X}\}$, where $t({\bf x})$ is the set of statistics in the neg-potential function $Q({\bf x})$ from the joint distribution of the node states.

This document will explore the behavior of the $\mu$-parametrized parameters for various dimensional models.

## Explorer

```{r, echo=FALSE, message=FALSE}

library(plyr)
library(dplyr)
library(reshape2)

stats <- function(H, V, type="negative") {
  if(type == "binary") t <- data.frame(c(0,1))[,rep(1, H + V)]
  else if(type == "negative") t <- data.frame(c(-1,1))[,rep(1, H + V)]
  else if(type == "experiment") t <- data.frame(c(1,3))[,rep(1, H + V)]
  else stop("Type must be 'binary', 'negative', or 'experiment'")

  names(t) <- c(paste0("h", 1:H), paste0("v", 1:V))

  #all possibilities of statistics in (H + V)-space
  t.grid <- expand.grid(t)

  for(i in 1:H){
    for(j in (H+1):(H+V)){
      t.grid <- cbind(t.grid, t.grid[,i]*t.grid[,j])
      names(t.grid)[ncol(t.grid)] <- paste0("theta", i, j - H)
    }
  }

  return(t.grid)
}

expected_value <- function(stats, theta, normalized = TRUE) {
  if(normalized) result <- t(as.matrix(stats)) %*% exp(as.matrix(stats) %*% as.matrix(theta)) %*% diag(1/apply(exp(as.matrix(stats) %*% as.matrix(theta)), 2, sum))
  else result <- t(as.matrix(stats)) %*% exp(as.matrix(stats) %*% as.matrix(theta))

  rownames(result) <- paste0("exp_", rownames(result))
  return(result)
}

reactiveSvg <- function(outputId) 
{
  div(id = outputId, class="d3graph")
}

shinyApp(
  ui = fluidPage(
      tags$head(
        includeScript("http://d3js.org/d3.v3.js"),
        includeCSS("www/css/surface3d.css"),
        includeCSS("www/css/linechart.css"),
        includeScript("www/js/surface3d.js"),
        includeScript("www/js/plots.js"),
        includeCSS("www/readable.min.css") 
      ),
      sidebarLayout(
        sidebarPanel(sliderInput("numHidden", label = "Number of hidden nodes:", 
                                 min = 1, max = 10, value = 1, step = 1),
                     sliderInput("numVisible", label = "Number of visible nodes:", 
                                 min = 1, max = 10, value = 1, step = 1),
                     radioButtons("encoding", label="Encoding:", 
                                  choices = list("Binary (0,1)" = "binary", "Negative (-1,1)" = "negative"),
                                  inline = FALSE),
                     selectizeInput("variableParams", "Parameters to vary:", 
                                    choices = NULL, multiple = TRUE,
                                    options = list(maxItems = 2))
        ),
        mainPanel(reactiveSvg(outputId = "surfacePlot"),
                  fluidRow(column(6, h4("Fixed Values"), uiOutput("fixedVals")),
                           column(6, h4("Variable Ranges"), uiOutput("variableVals"))
                  )
        )        
      )      
    ),
  server = function(input, output, session){
      stat_vals <- reactive({
        stats(H = input$numHidden, V = input$numVisible, type = input$encoding)
      })

      observe({
        vals <- stat_vals()
        updateSelectizeInput(session, inputId = "variableParams", choices = names(vals))
      })

      output$fixedVals <- renderUI({ 
        vals <- stat_vals()

        if(length(input$variableParams) > 0) {
          ## add fixed values
          lapply(names(vals)[!(names(vals) %in% input$variableParams)], function(name) {
            numericInput(paste0("fixedVal_", name), label = paste0(name, ":"), value=NULL)
          })    
        }
      })

      output$variableVals <- renderUI({
        vals <- stat_vals()

        if(length(input$variableParams) > 0) {
          ## add ranges for variable
          lapply(names(vals)[names(vals) %in% input$variableParams], function(name) {
            sliderInput(paste0("variableRange_", name), label = paste0(name, ":"), value=c(-10,10),
                        min=-20, max=20)
          })  
        }
      })

      output$surfacePlot <- reactive({ 
        vals <- stat_vals()

        if(length(input$variableParams) > 0) {
          fixed_vals <- data.frame(do.call("rbind", lapply(names(vals)[!(names(vals) %in% input$variableParams)], function(z) c(z, input[[paste0("fixedVal_", z)]]))))
          variable_ranges <- data.frame(do.call("rbind", lapply(names(vals)[names(vals) %in% input$variableParams], function(z) c(z, input[[paste0("variableRange_", z)]]))))

          if(ncol(fixed_vals) == 2) {
            names(fixed_vals) <- c("param","val")
            names(variable_ranges) <- c("param", "min", "max")

            if(sum(complete.cases(fixed_vals)) == nrow(fixed_vals) & nrow(variable_ranges) > 0) {
              ranges <- data.frame(apply(variable_ranges, 1, function(x) seq(as.numeric(x["min"]), as.numeric(x["max"]), length.out = 40)))
              ranges.grid <- expand.grid(ranges)

              ranges.grid <- ranges.grid %>%
                cbind(matrix(rep(as.numeric(as.character(fixed_vals[,"val"])), nrow(ranges.grid)), nrow = nrow(ranges.grid), byrow = TRUE))
              names(ranges.grid) <- c(as.character(variable_ranges$param), as.character(fixed_vals$param))

              surf_plot <- nrow(variable_ranges) == 2

              exp.grid0 <- t(expected_value(vals[,names(ranges.grid)], t(ranges.grid)))
              exp.grid <- cbind(exp.grid0, data.frame(ranges.grid[, variable_ranges$param])) 
              names(exp.grid) <- c(colnames(exp.grid0), as.character(variable_ranges$param))

              surf_values <- melt(exp.grid, id.vars=as.character(variable_ranges$param)) %>%
                dlply(.(variable), function(x){
                  dcast(x, formula = formula(paste0(variable_ranges$param[1], "~", variable_ranges$param[2])), value.var = "value") %>%
                    select(-1)
                }) %>%          
                llply(function(x) {
                  names(x) <- 0:(ncol(x) - 1) 
                  return(x)
                })

              surf_values <- llply(surf_values, function(x) x*10)

              line_values <- melt(exp.grid, id.vars=as.character(variable_ranges$param)) %>%
                dlply(.(variable), function(x){
                  x %>% select(-variable) %>%
                    mutate_("x" = as.character(variable_ranges$param))
                })

              if(surf_plot) return(list(values=surf_values, surf_plot=surf_plot))
              else return(list(values=line_values, surf_plot=surf_plot))
            }
          }
        }    
      })  
    },
  options = list(width="100%", height=600)
  )
```