### 2015-03-02 Toy Example/expected value/degen_magnitude.Rmd
example for meeting

---
title: "Degeneracy and parameter magnitude"
author: "Andee Kaplan"
date: "March 1, 2015"
output:
  html_document:
    theme: readable
runtime: shiny
---

```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(ggvis)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)
library(shiny)
```

```{r data, message=FALSE, warning=FALSE, echo=FALSE}
load("results.RData")

## params ----------------------------------
n <- 100
r_center <- seq(0.2, 3, by = 0.2)
r_exp <- .05

plot_data <- function(res) {

  plot.data <- data.frame()
  for(i in 1:nrow(res)) {
    tmp <- res$samp[[i]]
    H <- res[i,]$H
    V <- res[i,]$V

    tmp %>% 
      data.frame() %>% 
      rowwise() %>% 
      mutate_(ss_ratio = paste0("(", paste(paste0(names(.)[(H + V + 1):(H + V + H*V)], "^2"), collapse = " + "), ")/sum(", paste(paste0(names(.)[1:(H+V)],"^2"), collapse = " + "), ")")) %>%    
      ungroup() -> ratio

    inner_join(ratio, res$outside[[i]] %>% ungroup()) %>%
      select(ss_ratio, near_hull) %>%
      mutate(H = H, V = V, n_param = H + V + H*V) %>%
      rbind(plot.data) -> plot.data
  }
  return(plot.data)
}

plot.dat <- lapply(res, plot_data)
names(plot.dat) <- gsub("[.]", "_", r_center)

do.call(rbind, plot.dat) %>%
  mutate(name = rownames(do.call(rbind, plot.dat))) %>%
  separate(name, into = c("multiplier", "num"), "[.]") %>% 
  mutate(multiplier = as.numeric(gsub("_", ".", multiplier))) %>%
  mutate(n_param_f = factor(n_param)) %>%
  group_by(n_param_f, multiplier) %>%
  summarise(frac_degen = sum(near_hull)/n()) -> ggvis.dat
```

```{r plots, message=FALSE, warning=FALSE, echo=FALSE, out.height=300}
multiplier <- reactiveValues(current = 0.2)

on_click <- function(data, ...) {
  if(nrow(data) > 0) multiplier$current <- data$multiplier
}

ggvis.subset <- reactive({
  do.call(rbind, plot.dat) %>%
    mutate(name = rownames(do.call(rbind, plot.dat))) %>%
    mutate(n_param_f = factor(n_param)) %>%
    separate(name, into = c("multiplier", "num"), "[.]") %>% 
    mutate(multiplier = as.numeric(gsub("_", ".", multiplier))) %>%
    filter(multiplier == multiplier$current) 
})

ggvis.dat %>%
  ggvis(~multiplier, ~frac_degen) %>%
  layer_points(fill = ~n_param_f) %>% 
  layer_lines(stroke = ~n_param_f) %>%
  add_axis("x", title = "Radius Multiplier") %>%
  add_axis("y", title = "Fraction Degenerate") %>%
  add_legend("fill", title = "# Parameters") %>%
  add_legend("stroke", title = "# Parameters") %>%
  handle_click(on_click)

```

```{r plots2, message=FALSE, warning=FALSE, echo=FALSE, out.height=300}

ggvis.subset %>%
  group_by(multiplier, n_param_f, near_hull) %>%
  summarise(num = n()) %>%
  ggvis(~n_param_f, ~num) %>%
  layer_bars(fill = ~as.character(near_hull)) %>%
  add_axis("x", title = "# Parameters") %>%
  add_axis("y", title = "Count") %>%
  add_legend("fill", title = "Within .05 of Hull")

# ggvis.subset %>%
#   ggvis(~n_param_f, ~ss_ratio) %>%
#   layer_points(fill = ~as.character(near_hull), opacity := 0.4, size := 200) %>%
#   add_axis("x", title = "# Parameters") %>%
#   add_axis("y", title = "Ratio of SS of cross terms to main effects") %>%
#   add_legend("fill", title = "Within 0.05 of Hull")

renderPlot(
  ggvis.subset() %>%
    ggplot() + 
    geom_boxplot(aes(n_param_f, ss_ratio, colour = near_hull)) +
    xlab("# Parameters") +
    ylab("Ratio of sum of squares of cross terms to main effects") +
    theme_bw() +
    scale_colour_discrete("Within .05 of hull")
)
```

#Code Appendix
```{r eval = FALSE}
## load libraries ---------------------------
library(dplyr)
library(tidyr)
library(ggplot2)

## background functions -------------------------------------
stats <- function(H, V, type="negative") {
  if(type == "binary") t <- data.frame(c(0,1))[,rep(1, H + V)]
  else if(type == "negative") t <- data.frame(c(-1,1))[,rep(1, H + V)]
  else stop("Type must be 'binary', 'negative'")

  names(t) <- c(paste0("h", 1:H), paste0("v", 1:V))

  #all possibilities of statistics in (H + V)-space
  t.grid <- expand.grid(t)

  for(i in 1:H){
    for(j in (H+1):(H+V)){
      t.grid <- cbind(t.grid, t.grid[,i]*t.grid[,j])
      names(t.grid)[ncol(t.grid)] <- paste0("theta", i, j - H)
    }
  }

  return(data.matrix(t.grid))
}

expected_value <- function(theta, stats, normalized = TRUE) {
  result <- crossprod(t(crossprod(stats, exp(crossprod(t(stats), theta)))), diag(1/apply(exp(crossprod(t(stats), theta)), 2, sum), nrow=length(apply(exp(crossprod(t(stats), theta)), 2, sum))))
  rownames(result) <- paste0("exp_", rownames(result))
  return(result)
}

sample_sphere <- function(stat, n, r = 1) {
  require(dplyr)
  require(tidyr)

  samp <- function(space, r) {
    vec <- rnorm(space)
    paste0(r*vec/sqrt(sum(vec^2)), collapse=",")
  }

  res <- data.frame(samp = 1:n) %>%
    group_by(samp) %>%
    mutate(vals = samp(ncol(stat), r)) %>%
    separate(col = vals, sep = ",", into = colnames(stat)) %>%
    select(-samp) %>%
    data.matrix() %>%
    t()

  return(res)  
}

## main function --------------------------------------------
find_prop <- function(g_theta, stat, r_exp) {
  g_theta %>%
    data.frame() %>%
    group_by_(.dots = c(colnames(g_theta))) %>%
    do(samp = sample_points_outside(as.numeric(.), stat, r_exp)$in_hull) %>%
    group_by_(.dots = c(colnames(g_theta))) %>%
    do(as.data.frame(sum(.$samp[[1]]) > 0)) -> is_outside # count of not in hull greater than zero => there exists a point outside hull for this model

  is_outside %>% 
    rename_(near_hull = as.name(names(is_outside)[ncol(is_outside)])) %>%
    mutate(r = r_exp) -> inside_outside

  return(inside_outside)
}

## additional useful functions -------------------------------------
sample_points_outside <- function(g_theta, stats, r, n = 100) {
  #g_theta is the expected value mapping of a point theta in R^H+V+H*V
  x0 <- data.frame(samp = 1:n) %>%
    mutate(vals = samp_exp(length(g_theta), g_theta, r)) %>%
    separate(col = vals, sep = ",", into = colnames(stats)) %>%
    select(-samp)

  x0$in_hull <- !apply(x0, 1, in_hull, stats)

  return(x0)
}

samp_exp <- function(space, center, r) {
  vec <- rnorm(space)
  paste0(r*vec/sqrt(sum(vec^2)) + center, collapse=",")
}

in_hull <- function(point, hull_points) {
  require(lpSolveAPI)
  P <- data.matrix(point)
  A <- t(data.matrix(hull_points))

  lp_obj <- make.lp(nrow = 0, ncol = ncol(A))
  sapply(1:nrow(A), function(x) add.constraint(lp_obj, A[x,], type = "=", rhs = P[x]))
  add.constraint(lp_obj, rep(1, ncol(A)), type = "=", rhs = 1)
  sapply(1:ncol(A), function(x) add.constraint(lp_obj, 1, type = ">=", rhs = 0, indices = x))

  return(solve(lp_obj) == 0)
}

## params ----------------------------------
n <- 100
r_center <- seq(0.2, 3, by = 0.2)
#r_center <- 1
r_exp <- .05

## function for running simulation -------------------------------
get_res <- function(n, r_center, r_exp) {
  ## perform functions 
  test_cases <- expand.grid(data.frame(1:4)[,rep(1,2)]) %>% 
    rename(H = X1.4, V = X1.4.1) %>%
    filter(H <= V) %>% #remove those cases with less visibles than hiddens. Is this necessary?
    mutate(n_param = H*V + H + V) %>%
    mutate(max_facets = (2^(H+V))^(floor(n_param/2))) %>%
    mutate(n = n, r = r_center*(H + V)) %>%
    group_by(H, V, n_param, n, r) %>%
    do(stat = stats(.$H, .$V, "negative")) %>%
    ungroup() %>%
    group_by(H, V, n_param, n, r) %>%
    do(samp = sample_sphere(.$stat[[1]], .$n, .$r))

  test_cases_stat <- expand.grid(data.frame(1:4)[,rep(1,2)]) %>% 
    rename(H = X1.4, V = X1.4.1) %>%
    filter(H <= V) %>% #remove those cases with less visibles than hiddens. Is this necessary?
    mutate(n_param = H*V + H + V) %>%
    mutate(max_facets = (2^(H+V))^(floor(n_param/2))) %>%
    mutate(n = n, r = r_center*(H + V)) %>%
    group_by(H, V, n_param, n, r) %>%
    do(stat = stats(.$H, .$V, "negative"))

  test_cases <- inner_join(test_cases, test_cases_stat)

  test_cases %>%
    group_by(H, V, n_param, n, r) %>%
    do(g_theta = t(rbind(.$samp[[1]], expected_value(theta = .$samp[[1]], stats = data.matrix(.$stat[[1]]))))) -> cases

  cases <- inner_join(cases, test_cases)

  cases %>%
    group_by(H, V, r) %>%
    do(outside = find_prop(.$g_theta[[1]] %>% data.frame() %>% select(starts_with("exp")), .$stat[[1]], r_exp)) -> tmp

  res <- inner_join(tmp, cases %>% group_by(H, V, r, n_param) %>% do(samp = .$g_theta[[1]])) 
}

## run simulations -------------------------------------
res <- lapply(r_center, function(r) get_res(n, r, r_exp))
save(res, file = "results.RData")

```