### 2015-11-18 4x4 model/docs/ok_model.Rmd
reworking structure

---
title: "OK Restricted Boltzmann Machine"
author: "Andee Kaplan"
date: "October 28, 2015"
output: pdf_document
---

```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
source("../sample_functs.R")

load("../written/sample.params.Rdata")
load("../written/sample_images.Rdata")

sample.params %>%
  ungroup() %>%
  filter(!near_hull) %>%
  select(starts_with("v"), starts_with("h"), starts_with("theta"), -H, -V) %>%
  gather(parameter, value, everything()) %>%
  kable()

params_good <- list(main_hidden = sample.params %>% ungroup() %>% filter(!near_hull) %>% select(starts_with("h"), -H) %>% data.matrix(),
                    main_visible = sample.params %>% ungroup() %>% filter(!near_hull) %>% select(starts_with("v"), -V) %>% data.matrix(),
                    interaction = sample.params %>% ungroup() %>% filter(!near_hull) %>% select(starts_with("theta")) %>% data.matrix() %>% matrix(4))

visible_distn(params = params_good) %>%
  kable()

possibles <- stats(4, 4)
theta <- sample.params %>%
  ungroup() %>%
  filter(!near_hull) %>%
  select(starts_with("v"), starts_with("h"), starts_with("theta"), -H, -V) %>%
  data.matrix() %>% t

sum(exp(possibles %*% theta))

```