### 2016-02-18 presentations/coda2016/coda2016_poster.tex
added kaiser method on the grid and some writing

%%big poster size with big fonts
\documentclass[extrafontsizes, 30pt]{memoir}
\usepackage[paperheight=31in,paperwidth=47in,margin=1in,heightrounded,showframe]{geometry}

%%background from pptx
\usepackage{background}
\backgroundsetup{
scale=1,
angle=0,
opacity=1,  %% adjust
contents={\includegraphics[width=\paperwidth,height=\paperheight]{images/coda2016_poster_background.pdf}}
}

%%put words where I want
\usepackage[absolute, overlay]{textpos}
\setlength{\TPHorizModule}{1in}
\setlength{\TPVertModule}{\TPHorizModule}

%%tikz setup
\usepackage{tikz, subfig, amsthm}
\usepackage{tikz-3dplot}
\usetikzlibrary{arrows, shapes, positioning}

%%other options
\definecolor{isublue}{RGB}{58,117,196}
\definecolor{isured}{RGB}{206,17,38}
\counterwithout{figure}{chapter}
\usepackage{amsmath}
\setlength{\parindent}{0cm}

\begin{document}

%%-----------------------------------------------
%%
%%Middle big box
%%
\begin{textblock}{18.01}(14.53, 6.71)
{\large \bfseries Restricted Boltzmann machine (RBM)}

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\input{images/rbm.tikz}}
  \caption{An example restricted Boltzmann machine (RBM), which consists of two layers, a hidden ($\mathcal{H}$) and a visible layer ($\mathcal{V}$), with no connections within a layer. Hidden nodes indicated by gray circles and the visible nodes indicated by white circles.}
  \label{fig:rbm}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{images/visibles.tikz}
  \caption{Visibles diagram.}
  \label{fig:visibles}
\end{figure}

~\\[-1cm]
{\bfseries Joint distribution} \\[.25cm]
Let $\boldsymbol x = \{h_1, \dots, h_H, v_1,\dots,v_V\}$ represent the states of the visible and hidden nodes in an RBM. Then  the probability corresponding to the state of each node taking the value of $1$:
\begin{align}
\label{eqn:pmf}
f_{\boldsymbol \theta} (\boldsymbol x) = \dfrac{\exp\left(\sum\limits_{i = 1}^V \sum\limits_{j=1}^H \theta_{ij} v_i h_j + \sum\limits_{i = 1}^V\theta_{v_i} v_i + \sum\limits_{j = 1}^H\theta_{h_j} h_j\right)}{\sum\limits_{\boldsymbol x \in \mathcal{X}}\exp\left(\sum\limits_{i = 1}^V \sum\limits_{j=1}^H \theta_{ij} v_i h_j + \sum\limits_{i = 1}^V\theta_{v_i} v_i + \sum\limits_{j = 1}^H\theta_{h_j} h_j\right)} 
\end{align}
\end{textblock}

%%-----------------------------------------------
%%
%%Top left box
%%
\begin{textblock}{14.08}(.8, 5.71)
{\large \bfseries Deep learning}

\end{textblock}

%%-----------------------------------------------
%%
%%Middle left box
%%
\begin{textblock}{14.08}(.8, 11.83)
{\large \bfseries Degeneracy in random graph models}

\end{textblock}

%%-----------------------------------------------
%%
%%Bottom left box
%%
\begin{textblock}{14.08}(.8, 20.64)
{\large \bfseries Tiny example}

\end{textblock}

%%-----------------------------------------------
%%
%%Top right box
%%
\begin{textblock}{14.08}(33.5, 5.71)
{\large \bfseries Managable examples}
\end{textblock}

%%-----------------------------------------------
%%
%%Middle right box
%%
\begin{textblock}{14.08}(33.5, 12.91)
{\large \bfseries Fitting}

\end{textblock}

%%-----------------------------------------------
%%
%%Bottom right box
%%
\begin{textblock}{14.08}(33.5, 23.85)
{\large \bfseries Discussion}
\end{textblock}

%%-----------------------------------------------
%%
%%Bottom middle box
%%
\begin{textblock}{15.08}(16.17, 24.84)
{\large \bfseries References}
\end{textblock}

\end{document}