### 2016-10-01 writing/appendix-instability.Rmd
body of instability rewrite complete

\begin{align*}
&\max\left\{\log\frac{P_{\boldsymbol \theta}(v_1, \dots, v_V)}{P_{\boldsymbol \theta}(v^*_1, \dots, v^*_V)}: (v_1, \dots, v_V) \text{ and } (v^*_1, \dots, v^*_V) \in \{-1, 1\}^V \text{ differ by one component} \right\} \\
&\quad = \max\limits_{k \in 1, \dots, V}\left\{\log\frac{\sum\limits_{\boldsymbol h \in \{-1,1\}^H} \exp\left(
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right)}{\sum\limits_{\boldsymbol h \in \{-1,1\}^H} \exp\left(
\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right)}: v_k \not= v^*_k \right\} \\
& \quad \ge \max\limits_{k \in 1, \dots, V}\left\{\log\frac{\max\limits_{\boldsymbol h \in \{-1,1\}^H}\exp\left(
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right)}{2^H\max\limits_{\boldsymbol h \in \{-1,1\}^H}\exp\left(
\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right)}: v_k \not= v^*_k \right\} \\
&\quad = \max\limits_{k \in 1, \dots, V}\left\{\max\limits_{\boldsymbol h \in \{-1,1\}^H}\left\{
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right\}\right. \\
& \quad \quad \left. - \max\limits_{\boldsymbol h \in \{-1,1\}^H}
\left\{\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right\} - H\log2: v_k \not= v^*_k \right\} \\
&\quad \ge \max\limits_{\boldsymbol v \in \{-1,1\}^V}\max\limits_{\boldsymbol h \in \{-1,1\}^H}Q(\boldsymbol x) - \min\limits_{\boldsymbol v \in \{-1,1\}^V}\max\limits_{\boldsymbol h \in \{-1,1\}^H}Q(\boldsymbol x) - H\log 2
\end{align*}