### 2016-10-01 writing/appendix-instability.Rmd
add proofs for propositions

{\bf Proof of Proposition \ref{prop:instab}.} We prove the contrapositive. Suppose that $\Delta_V(\boldsymbol \theta) \le C$ holds for some $C > 0$. Under the RBM model fo visibles, $P_{\boldsymbol \theta}(\boldsymbol v) > 0$ holds for each outcome $\boldsymbol v \in \{-1,1\}^V$. Let $\boldsymbol v_{min} \equiv \argmin\limits_{\boldsymbol v \in \{-1, 1\}^V}P_{\boldsymbol \theta}(\boldsymbol v)$ and $\boldsymbol v_{max} \equiv \argmax\limits_{\boldsymbol v \in \{-1, 1\}^V}P_{\boldsymbol \theta}(\boldsymbol v)$. Note there exists a sequence $\boldsymbol v_{min} \equiv \boldsymbol v_0, \boldsymbol v_1, \dots, \boldsymbol v_k \equiv \boldsymbol v_{max}$ in $\{-1,1\}^V$ of component-wise switches to move from $\boldsymbol v_{min}$ to $\boldsymbol v_{max}$ in the sample space (i.e. $\boldsymbol v_i, \boldsymbol v_{i + 1} \in \{-1,1\}^V$ differ by exactly $1$ component for $i = 0, \dots, k -1$) for some integer $k \in \{0, 1, \dots, V\}$. Then
\begin{align*}
\log\left[\frac{P_{\boldsymbol \theta}(\boldsymbol v_{max})}{P_{\boldsymbol \theta}(\boldsymbol v_{min})}\right] &= \left|\sum\limits_{i = 1}^k\log\left(\frac{P_{\boldsymbol \theta}(\boldsymbol v_i)}{P_{\boldsymbol \theta}(\boldsymbol v_{i-1})}\right)\right| \\
& \le \sum\limits_{i = 1}^k\left|\log\left(\frac{P_{\boldsymbol \theta}(\boldsymbol v_i)}{P_{\boldsymbol \theta}(\boldsymbol v_{i-1})}\right)\right| \\
& \le k \Delta_V(\boldsymbol \theta) \le VC
\end{align*}
using $k \le V$ and $\Delta_V(\boldsymbol \theta) \le C$.$\Box$

{\bf Proof of Proposition \ref{prop:degen}.} Define $\boldsymbol v_{max}$ and $\boldsymbol v_{min}$ as in the proof of Proposition \ref{prop:instab}. Fix $0 < \epsilon < 1$ Then, $\boldsymbol x_{max} \in M_{\epsilon, \boldsymbol \theta}$, so $P_{\boldsymbol \theta}(M_{\epsilon, \boldsymbol \theta}) \ge P_{\boldsymbol \theta}(\boldsymbol v_{max})$. If $\boldsymbol v \in \{-1, 1\}^V \setminus M_{\epsilon, \boldsymbol \theta}$, then by definition $P_{\boldsymbol \theta}(\boldsymbol v) \le [P_{\boldsymbol \theta}(\boldsymbol v_{max})]^{1-\epsilon}[P_{\boldsymbol \theta}(\boldsymbol v_{min})]^{\epsilon}$ holds so that
\begin{align*}
1-P_{\boldsymbol \theta}(M_{\epsilon, \boldsymbol \theta}) &= P_{\boldsymbol \theta}(M_{\epsilon, \boldsymbol \theta}^C) \\
& = \sum\limits_{\boldsymbol v \in \{-1, 1\}^V \setminus M_{\epsilon, \boldsymbol \theta}}P_{\boldsymbol \theta}(\boldsymbol v) \\
& \le (2^V)[P_{\boldsymbol \theta}(\boldsymbol v_{max})]^{1-\epsilon}[P_{\boldsymbol \theta}(\boldsymbol v_{min})]^{\epsilon}
\end{align*}
Then,
\begin{align*}
\frac{1}{V}\log\left[\frac{P_{\boldsymbol \theta}(M_{\epsilon, \boldsymbol \theta})}{1-P_{\boldsymbol \theta}(M_{\epsilon, \boldsymbol \theta})}\right] & \ge \frac{1}{V} \log\left[\frac{P_{\boldsymbol\theta}(\boldsymbol v_{max})}{(2^V)[P_{\boldsymbol \theta}(\boldsymbol v_{max})]^{1-\epsilon}[P_{\boldsymbol \theta}(\boldsymbol v_{min})]^{\epsilon}}\right] \\
&= \frac{\epsilon}{V} \log\left[\frac{P_{\boldsymbol \theta}(\boldsymbol v_{max})}{P_{\boldsymbol \theta}(\boldsymbol v_{min})}\right] - \log 2 \rightarrow \infty
\end{align*}
as $V \rightarrow \infty$ by the definition of an unstable RBM model.$\Box$

\begin{align*}
&\max\left\{\log\frac{P_{\boldsymbol \theta}(v_1, \dots, v_V)}{P_{\boldsymbol \theta}(v^*_1, \dots, v^*_V)}: (v_1, \dots, v_V) \text{ and } (v^*_1, \dots, v^*_V) \in \{-1, 1\}^V \text{ differ by one component} \right\} \\
&\quad = \max\limits_{k \in 1, \dots, V}\left\{\log\frac{\sum\limits_{\boldsymbol h \in \{-1,1\}^H} \exp\left(
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right)}{\sum\limits_{\boldsymbol h \in \{-1,1\}^H} \exp\left(
\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right)}: v_k \not= v^*_k \right\} \\
& \quad \ge \max\limits_{k \in 1, \dots, V}\left\{\log\frac{\max\limits_{\boldsymbol h \in \{-1,1\}^H}\exp\left(
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right)}{2^H\max\limits_{\boldsymbol h \in \{-1,1\}^H}\exp\left(
\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right)}: v_k \not= v^*_k \right\} \\
&\quad = \max\limits_{k \in 1, \dots, V}\left\{\max\limits_{\boldsymbol h \in \{-1,1\}^H}\left\{
\sum\limits_{i = 1}^V \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i = 1}^V\sum\limits_{j = 1}^H \theta_{ij}v_i h_j\right\}\right. \\
& \quad \quad \left. - \max\limits_{\boldsymbol h \in \{-1,1\}^H}
\left\{\sum\limits_{i \not= k} \theta_{v_i}v_i + \sum\limits_{j = 1}^H \theta_{h_j}h_j + \sum\limits_{i \not=k}\sum\limits_{j = 1}^H \theta_{ij}v_i h_j + \theta_{v_k}v_k^* + v_k^*\sum\limits_{j = 1}^H \theta_{kj}h_j
\right\} - H\log2: v_k \not= v^*_k \right\} \\
&\quad \ge \max\limits_{\boldsymbol v \in \{-1,1\}^V}\max\limits_{\boldsymbol h \in \{-1,1\}^H}Q(\boldsymbol x) - \min\limits_{\boldsymbol v \in \{-1,1\}^V}\max\limits_{\boldsymbol h \in \{-1,1\}^H}Q(\boldsymbol x) - H\log 2
\end{align*}