### 2016-12-02 code/degeneracy/apps/mu_parameterization_3d.Rmd

---
output:
  html_document:
    theme: readable
runtime: shiny
---

# $\mu$ parameterization in 3 dimensions

For one hidden and one visible node, we sample points uniformly on a 3 dimensional sphere of radius `r` centered at `(0,0,0)` and view the resulting expected values in 3 dimensions.
```{r, echo=FALSE, message=FALSE}
source("functions.R")
library(scatterplot3d)
library(plyr)
library(dplyr)
library(reshape2)
library(ggvis)
library(shinythemes)

## Num nodes
H <- 1
V <- 1

## Num to sample
n <- 500

## Radius to sample
r <- c(seq(0,1, by=0.05), seq(1.5, 10, by=0.5))

```

```{r, echo=FALSE, message=FALSE}

shinyApp(
  ui=
    fluidPage(
      theme = shinytheme("readable"),
      fluidRow(style = "padding-bottom: 20px;",
        column(2, radioButtons("encoding", label="Encoding:", 
                                      choices = list("Negative (-1,1)" = "negative", "Binary (0,1)" = "binary"),
                                      inline = FALSE)),
        column(4, sliderInput("rSlider", label = "Radius:", min = 0, max = 10, value=0, step = 0.1, animate = list(interval=1000, loop=FALSE, playButton=icon("play"), pauseButton=icon("pause")))),
        column(6, 
               column(3, span("Translate origin by (x, y, z): ")),
               column(3, numericInput("x", value=NA, label="")),
               column(3, numericInput("y", value=NA, label="")), 
               column(3, numericInput("z", value=NA, label=""))
               )
      ),
      fluidRow(
        column(6, plotOutput("exp_val_3d")),
        column(6, ggvisOutput("min_dist_hist"))
      )
    ),
  server=function(input, output, session){

    data <- reactive({
      x <- ifelse(is.na(input$x), 0, input$x)
      y <- ifelse(is.na(input$y), 0, input$y)
      z <- ifelse(is.na(input$z), 0, input$z)

      stat <- stats(H, V, input$encoding)
      theta <- sample_sphere(stat, n, input$rSlider) + matrix(rep(c(x,y,z), n), nrow = 3)
      exp_vals <- expected_value(stats = stat, theta = theta, normalized = TRUE)

      return(list(stat=stat, exp_vals=exp_vals))
    })

#     cdf_mu <- reactive({
#       dat <- data()
#       stat <- dat$stat
#       exp_vals <- dat$exp_vals
#       
#       cdf_mu <- melt(apply(apply(t(exp_vals), 2, abs), 2, sort)) %>%
#         group_by(Var2)
#       cdf_mu$Var1 <- cdf_mu$Var1/ncol(exp_vals)
#       
#       return(cdf_mu)
#     })

    min_dist_data <- reactive({
      dat <- data()
      stat <- dat$stat
      exp_vals <- dat$exp_vals

      lines <- generate_hull_from_stats(stat)
      dist <- do.call(rbind, apply(exp_vals, 2, function(x){ min_distance_to_hull(x, lines) }))
      return(dist)      
    })

    output$exp_val_3d <- renderPlot({
      dat <- data()
      stat <- dat$stat
      exp_vals <- dat$exp_vals

      s <- scatterplot3d(stat, pch='.')

      segs <- NULL
      for(i in 1:nrow(stat)) {
        for(j in 1:nrow(stat)) {
          if(i < j) {
            start <- s$xyz.convert(t(stat[i,]))
            end <- s$xyz.convert(t(stat[j,]))
            segs <- rbind(segs, c(start$x, start$y, end$x, end$y))
          }
        }
      }

      segments(segs[,1], segs[,2], segs[,3], segs[,4], lty=2)
      points(s$xyz.convert(t(exp_vals)), col="blue", pch=19)

    })

#     cdf_mu %>%
#       ggvis(~value, ~Var1, stroke = ~Var2) %>%
#       layer_paths() %>%
#       scale_numeric("x", domain = c(0,1)) %>%
#       scale_numeric("y", domain = c(0,1)) %>%
#       add_axis("x", title = "Absolute Expected Value") %>%
#       add_axis("y", title = "Empirical CDF") %>%
#       add_legend("stroke", title = "Mu Parameterized") %>%
#       set_options(width=200, height=200, keep_aspect = TRUE, resizable = FALSE) %>%
#       bind_shiny(plot_id = "exp_val_ecdf")

    min_dist_data %>%
      ggvis(~min_dist) %>%
      layer_histograms() %>%
      scale_numeric("x", domain = c(0,1)) %>%
      add_axis("x", title = "Minimum approximate distance to convex hull edge") %>%
      set_options(width="auto", height=100, keep_aspect = TRUE, resizable = FALSE) %>%
      bind_shiny(plot_id = "min_dist_hist")
  },
  options = list(height = 600)
)

```