### 2017-11-02 writing/paper-note/appendices/appendix-note.Rmd
forgot to push last time

\clearpage

# (APPENDIX) Appendix {-}

# Proofs of instability results {#appendix-instab}

**Proof of Proposition \@ref(thm:instab-elpr).** For part (i), we prove the contrapositive, supposing that $\DN(\boldsymbol \theta_N) \le C$ holds for some $C > 0$ and show $\REP(\boldsymbol \theta_N) \leq NC$. Let $\boldsymbol x_{min} \equiv \argmin\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$
and $\boldsymbol x_{max} \equiv \argmax\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$. Note there exists a sequence $\boldsymbol x_{min} \equiv \boldsymbol x_0, \boldsymbol x_1, \dots, \boldsymbol x_k \equiv \boldsymbol x_{max}$ in $\mathcal{X}^N$ of component-wise switches to move from $\boldsymbol x_{min}$ to $\boldsymbol x_{max}$ in the sample space (i.e. $\boldsymbol x_i, \boldsymbol x_{i + 1} \in \mathcal{X}^N$ differ in exactly $1$ component, $i = 0, \dots, k$) for some integer $k \in \{0, 1, \dots, N\}$. Under the FOES model, recall
$P_{\boldsymbol \theta_N}(\boldsymbol x) > 0$ holds so that $\log P_{\boldsymbol \theta_N}(\boldsymbol x)$ is well-defined for each outcome $\boldsymbol x \in \mathcal{X}^N$. Then, if $k > 0$, it follows that
\begin{align*}
\REP(\boldsymbol \theta_N) = \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] &= \left|\sum\limits_{i = 1}^k\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta_N}(\boldsymbol x_{i-1})}\right)\right| \\
&\le \sum\limits_{i = 1}^k\left|\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta}(\boldsymbol x_{i-1})}\right)\right| \le k \Delta_N(\boldsymbol \theta_N) \le NC,
\end{align*}
using $k \le N$ and $\Delta(\boldsymbol \theta_N) \le C$. If $k = 0$, then $\boldsymbol x_{max} = \boldsymbol x_{min}$ and the same bound above holds. This establishes part (i). To show part (ii), note the definition of S-instability (i.e., $\lim_{N\to \infty}\REP(\boldsymbol \theta_N)/N= \infty$) combined with part (i) implies that $\lim_{N\to \infty}\DN(\boldsymbol \theta_N)=\infty$. \hfill $\Box$

**Proof of Proposition \@ref(thm:degenFOES).** As $|\mathcal{X}|<\infty$ holds in the FOES model, we may suppose
$|\mathcal{X}|>1$; otherwise, $\mathcal{X}^N$ has one outcome and the model is  trivially degenerate for all $N \geq 1$. Fix $0 < \epsilon < 1$ and write $\boldsymbol x_{min} \equiv \argmin\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$ and $\boldsymbol x_{max} \equiv \argmax\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$. Then, $\boldsymbol x_{max} \in M_{\epsilon, \boldsymbol \theta_N}$, so $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \ge P_{\boldsymbol \theta_N}(\boldsymbol x_{max}) > 0$. If $\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}$, then by definition
$P_{\boldsymbol \theta_N}(\boldsymbol x) \le [P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}$ holds so that
$$
1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})
 = \sum\limits_{\boldsymbol x \in \mathcal{X}^N \setminus M_{\epsilon, \boldsymbol \theta_N}}P_{\boldsymbol \theta_N}(\boldsymbol x)
  \le (|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}.
$$
From the lower bound on $P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})$ and the
upper bound on $1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})$, it follows that
\begin{align*}
\frac{1}{N}\log\left[\frac{P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}{1-P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N})}\right] & \ge \frac{1}{N} \log\left[\frac{P_{\boldsymbol\theta_N}(\boldsymbol x_{max})}{(|\mathcal{X}|^N)[P_{\boldsymbol \theta_N}(\boldsymbol x_{max})]^{1-\epsilon}[P_{\boldsymbol \theta_N}(\boldsymbol x_{min})]^{\epsilon}}\right] \\
&= \frac{\epsilon}{N} \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] - \log |\mathcal{X}| \rightarrow \infty
\end{align*}
as $N \rightarrow \infty$ by the definition of an S-unstable FOES model \@ref(eq:Sun). Consequently,
$P_{\boldsymbol \theta_N}(M_{\epsilon, \boldsymbol \theta_N}) \rightarrow 1$ as $N \rightarrow \infty$ as claimed. \hfill $\Box$

**Proof of Corollary \@ref(cor:sign).** The model condition PSR implies that
\begin{equation}
(\#eq:R)
\frac{\max_{\boldsymbol y \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol y)  }{\min_{\boldsymbol y \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol y) } = \frac{\max_{\boldsymbol y \in \mathcal{X}^N}P_{-\boldsymbol \theta_N}(\boldsymbol y) }{\min_{\boldsymbol y \in \mathcal{X}^N}P_{-\boldsymbol \theta_N}(\boldsymbol y) }
\end{equation}
so that the log-ratio $\REP(\boldsymbol \theta_N)=\REP(-\boldsymbol \theta_N)$ is the same for both $\boldsymbol \theta_N$ and $-\boldsymbol \theta_N$ in \@ref(eq:elpr). Now part (i) of Corollary \@ref(cor:sign) follows from $\REP(\boldsymbol \theta_N)/N=\REP(-\boldsymbol \theta_N)/N\to \infty$ as $N\to \infty$ in \@ref(eq:Sun). To show part (ii), fix $0 < \epsilon < 1$ and consider a $\epsilon$-mode set $\mathcal{M}_{\epsilon, \boldsymbol \theta_N}$ under $\boldsymbol \theta_N$ from \@ref(eq:mode). If $\boldsymbol x \in \mathcal{M}_{\epsilon, \boldsymbol \theta_N}^c \equiv \mathcal{X}^N \setminus \mathcal{M}_{\epsilon, \boldsymbol \theta_N}$, then, by definition,
$$
\frac{P_{\boldsymbol \theta_N}(\boldsymbol x)}{\min_{\boldsymbol y \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol y)} \leq \left[\frac{\max_{\boldsymbol y \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol y)  }{\min_{\boldsymbol y \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol y) } \right]^{1-\epsilon}
$$
holds, which is equivalent to
$$
\frac{\max_{\boldsymbol y \in \mathcal{X}^N}P_{-\boldsymbol \theta_N}(\boldsymbol y)}{P_{-\boldsymbol \theta_N}(\boldsymbol x)} \leq
\left[\frac{\max_{\boldsymbol y \in \mathcal{X}^N}P_{-\boldsymbol \theta_N}(\boldsymbol y)  }{\min_{\boldsymbol y \in \mathcal{X}^N}P_{-\boldsymbol \theta_N}(\boldsymbol y) } \right]^{1-\epsilon}
$$
by model condition PSR and \@ref(eq:R). The latter is in turn equivalent to
\begin{equation}
(\#eq:R2)
\log P_{-\boldsymbol \theta_N}(\boldsymbol x) \geq \epsilon \max\limits_{\boldsymbol y \in \mathcal{X}^N} \log  P_{-\boldsymbol \theta_N}(\boldsymbol y) + (1-\epsilon)\min\limits_{\boldsymbol y \in \mathcal{X}^N} \log P_{-\boldsymbol \theta_N}(\boldsymbol y),
\end{equation}
so that $\boldsymbol x \in \mathcal{M}_{\epsilon, \boldsymbol \theta_N}^c$ if and only if \@ref(eq:R2) holds. Next consider the $(1-\epsilon)$-mode set $\mathcal{M}_{1-\epsilon, -\boldsymbol \theta_N}$ under $-\boldsymbol \theta_N$ from \@ref(eq:mode). If $\boldsymbol x \in\mathcal{M}_{1-\epsilon, -\boldsymbol \theta_N}$, then by definition $\boldsymbol x$ fulfills \@ref(eq:R2) and so $\boldsymbol x \in \mathcal{M}_{\epsilon, \boldsymbol \theta_N}^c$, showing that $\mathcal{M}_{1-\epsilon, -\boldsymbol \theta_N} \subset \mathcal{M}_{\epsilon, \boldsymbol \theta_N}^c$. By this and the fact that that Theorem \@ref(thm:degenFOES) and Corollary \@ref(cor:sign)(i) entail that $P_{-\boldsymbol \theta_N}(\boldsymbol X_N \in \mathcal{M}_{1-\epsilon, -\boldsymbol \theta_N})\to 1$ as $N\to \infty$ (i.e., $P_{-\boldsymbol \theta_N}$ is S-unstable), we have
$$
1  = \lim_{N\to \infty} P_{-\boldsymbol \theta_N}(\boldsymbol X_N \in \mathcal{M}_{1-\epsilon, -\boldsymbol \theta_N}) \leq \lim_{N\to \infty} P_{-\boldsymbol \theta_N}(\boldsymbol X_N \in \mathcal{M}_{\epsilon, \boldsymbol \theta_N}^c ) \leq 1,
$$
proving Corollary \@ref(cor:sign)(ii) \hfill $\Box$

**Proof of Proposition \@ref(prp:prop1).** For any two outcomes $\boldsymbol x_1, \boldsymbol x_2\in\mathcal{X}^N$, the log-ratio of probabilities from the linear exponential model \@ref(eq:expo) with $k$ parameters/statistics satisfies
$$
\left|\log \left[ \frac{P_{\boldsymbol \theta_N}(\boldsymbol x_1)}{P_{\boldsymbol \theta_N}(\boldsymbol x_2)}  \right] \right| =
\left|  \sum_{i=1}^k \theta_{i,N} [g_{i,k}(\boldsymbol x_1) - g_{i,k}(\boldsymbol x_2) ] \right|  \leq  \sum_{i=1}^k | \theta_{i,N}| (U_{i,N}-L_{i,N});
$$
consequently, $\REP(\boldsymbol \theta_N )  \leq  \sum_{i=1}^k | \theta_{i,N}| (U_{i,N}-L_{i,N})$ holds in \@ref(eq:elpr) and model stability in Proposition \@ref(prp:prop1) follows from \@ref(eq:Sun). \hfill $\Box$

**Proof of Proposition \@ref(prp:prop2).**
Writing $\boldsymbol x=(x_1,\ldots,x_N)$ and $\boldsymbol h = (h_1,\ldots,h_{N_{\mathcal{H}}})$ with all components $x_i,h_j\in\{\pm 1\}$, probabilities in the joint RBM model \@ref(eq:RBM1) can be written as $\tilde{P}_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h) = c(\boldsymbol \theta_N)\exp[ f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)]$ in terms of the function $f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)$ from \@ref(eq:f) and the normalizing constant $c(\boldsymbol \theta_N)= \exp [-\psi(\boldsymbol \theta_N)]$ from \@ref(eq:RBM1). Let $\boldsymbol x_M, \boldsymbol x_m\in\{\pm 1\}^N$ be such that $P_{\boldsymbol \theta_N} (\boldsymbol x_M) = \max_{\boldsymbol x}P_{\boldsymbol \theta_N} (\boldsymbol x)$ and $P_{\boldsymbol \theta_N} (\boldsymbol x_m) = \min_{\boldsymbol x}P_{\boldsymbol \theta_N} (\boldsymbol x)$ under the marginal RBM model $P_{\boldsymbol \theta_N} (\boldsymbol x) = c(\boldsymbol \theta_N)\sum_{\boldsymbol h \in\{\pm 1\}^{\mathcal{N}_H}} \tilde{P}_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)= c(\boldsymbol \theta_N)\sum_{\boldsymbol h \in\{\pm 1\}^{\mathcal{N}_H}} \exp[ f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)]$ from \@ref(eq:RBM2). Also, $\boldsymbol x_0,x_1\in\{\pm 1\}^N$ be such that $\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x_0, \boldsymbol h)=\max_{\boldsymbol x}\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h)$ and $\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x_1, \boldsymbol h)=\min_{\boldsymbol x}\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h)$. Then, Proposition \@ref(prp:prop2)(i) follows from $\REP_{(\boldsymbol X)}(\boldsymbol \theta_N) = \log[P_{\boldsymbol \theta_N} (\boldsymbol x_M) /P_{\boldsymbol \theta_N} (\boldsymbol x_m) ]$ and the lower/upper bounds on $P_{\boldsymbol \theta_N} (\boldsymbol x_M)$ and $P_{\boldsymbol \theta_N} (\boldsymbol x_m)$ as
$$
c(\boldsymbol \theta_N) \exp[\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x_0 , \boldsymbol h)]
\leq P_{\boldsymbol \theta_N} (\boldsymbol x_0) \leq  P_{\boldsymbol \theta_N} (\boldsymbol x_M) \leq  2^{N_{\mathcal{H}}} c(\boldsymbol \theta_N) \exp[\max_{\boldsymbol x}\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)]
$$
and
\begin{align*}
c(\boldsymbol \theta_N) \exp[\min_{\boldsymbol x}\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h)] &\leq c(\boldsymbol \theta_N) \exp[ \max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x_m , \boldsymbol h)] \\
&\leq P_{\boldsymbol \theta_N} (\boldsymbol x_m) \\
&\leq P_{\boldsymbol \theta_N} (\boldsymbol x_1) \\ 
&\leq 2^{N_{\mathcal{H}}} c(\boldsymbol \theta_N) \exp[ \max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x_1 , \boldsymbol h)]\\&=&2^{N_{\mathcal{H}}} c(\boldsymbol \theta_N) \exp[\min_{\boldsymbol x}\max_{\boldsymbol h}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h)]
\end{align*}

To prove Proposition \@ref(prp:prop2), we next expand the function $f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)$ from \@ref(eq:f) as
$$
f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h)= \sum_{j=1}^{N_\mathcal{H}} h_j \theta_{j,N}^{\mathcal{H}}  +   \sum_{i=1}^N \left( \theta_{i,N}^{\mathcal{V}} + \sum_{j=1}^{N_\mathcal{H}}  h_j  \theta_{ij,N}^{\mathcal{VH}}\right)x_i=\sum_{i=1}^N x_i  \theta_{i,N}^{\mathcal{V}} +  \sum_{j=1}^{N_\mathcal{H}} \left( \theta_{j,N}^{\mathcal{H}}  + \sum_{i=1}^N  x_i  \theta_{ij,N}^{\mathcal{VH}}\right)h_j.
$$
By this and the fact that $x_i,h_j\in\{\pm 1\}$, we then have
\begin{align}
\nonumber \max_{\boldsymbol x}  f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h) =  \sum_{j=1}^{N_\mathcal{H}} h_j \theta_{j,N}^{\mathcal{H}}  +   a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h), & \min_{\boldsymbol x}  f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h) =  \sum_{j=1}^{N_\mathcal{H}} h_j \theta_{j,N}^{\mathcal{H}}  - a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h),\\
\nonumber \max_{\boldsymbol h}  f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h) = \sum_{i=1}^{N}x_i \theta_{i,N}^{\mathcal{V}} + b_{\boldsymbol \theta_N, \mathcal{V}} (\boldsymbol x), & \min_{\boldsymbol h}  f_{\boldsymbol \theta_N} (\boldsymbol x, \boldsymbol h) = \sum_{i=1}^{N}x_i \theta_{i,N}^{\mathcal{V}}  -   b_{\boldsymbol \theta_N, \mathcal{V}} (\boldsymbol x), \\ 
(\#eq:max)
a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h) \equiv \sum_{i=1}^N \left| \theta_{i,N}^{\mathcal{V}} + \sum_{j=1}^{N_\mathcal{H}}  h_j  \theta_{ij,N}^{\mathcal{VH}}\right|, &  b_{\boldsymbol \theta_N, \mathcal{V}} (\boldsymbol x) \equiv \sum_{j=1}^{N_\mathcal{H}} \left| \theta_{j,N}^{\mathcal{H}}  + \sum_{i=1}^N  x_i  \theta_{ij,N}^{\mathcal{VH}}\right|,
\end{align}
where  $\boldsymbol h^T \boldsymbol \theta_N^{\mathcal{H}}=\sum_{j=1}^{N_\mathcal{H}} h_j \theta_{j,N}^{\mathcal{H}}$, $\boldsymbol x^T \boldsymbol \theta_N^{\mathcal{V}}=  \sum_{i=1}^{N}x_i \theta_{i,N}^{\mathcal{V}}$ and $\Gam\equiv \max_{\boldsymbol h} a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h)$.
From this, it follows that
\begin{align*}
\REP_{(\boldsymbol X, \boldsymbol H)}(\boldsymbol \theta_N) &=  \max_{\boldsymbol h}\max_{\boldsymbol x}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h) -  \min_{\boldsymbol h}\min_{\boldsymbol x}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h)\\
&= \max_{\boldsymbol h_1 }  \max_{\boldsymbol h_2 }\left[ (\boldsymbol h_1 - \boldsymbol h_2)^T \boldsymbol \theta_N^{\mathcal{H}}  +   a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_1)  + a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_2)\right],
\end{align*}
which leads to the upper bound $\REP_{(\boldsymbol X, \boldsymbol H)}(\boldsymbol \theta_N) \leq 2 \Gam + 2 |\boldsymbol \theta_N^{\mathcal{H}} |_1$. Then, taking $\boldsymbol h_1=\boldsymbol h_2$ (i.e., before maximization) gives $\REP_{(\boldsymbol X, \boldsymbol H)}(\boldsymbol \theta_N) \geq 2\Gam$ and taking $\boldsymbol h_1=-\boldsymbol h_2$, such that $\boldsymbol h_1^T \boldsymbol \theta_N^{\mathcal{H}} = |\theta_N^{\mathcal{H}}|_1$, gives $\REP_{(\boldsymbol X, \boldsymbol H)}(\boldsymbol \theta_N) \geq 2|\theta_N^{\mathcal{H}}|_1$; this yields the lower bound $\REP_{(\boldsymbol X, \boldsymbol H)}(\boldsymbol \theta_N)\geq 2\max\{\Gam, |\boldsymbol \theta_N^{\mathcal{H}} |_1\}$.

We next consider $\elt$ and, by \@ref(eq:max), write
\begin{align*}
\elt &=  \max_{\boldsymbol h}\max_{\boldsymbol x}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h) -  \max_{\boldsymbol h}\min_{\boldsymbol x}f_{\boldsymbol \theta_N} (\boldsymbol x , \boldsymbol h)\\
&= \max_{\boldsymbol h_1} \min_{\boldsymbol h_2 }\left[ (\boldsymbol h_1 - \boldsymbol h_2)^T \boldsymbol \theta_N^{\mathcal{H}} + a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_1)  + a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_2)\right].
\end{align*}
Taking $\boldsymbol h_1=\boldsymbol h_2$ and maximizing over both $\boldsymbol h_1,\boldsymbol h_2$ produces the upper bound $\elt \leq 2\Gam$. Then, using $(\boldsymbol h_1 - \boldsymbol h_2)^T \boldsymbol \theta_N^{\mathcal{H}} + a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_2) \geq - 2|\boldsymbol \theta_N^{\mathcal{H}} |_1$ and maximizing over $\boldsymbol h_1$ gives $\elt \geq  \Gam- 2|\boldsymbol \theta_N^{\mathcal{H}} |_1$, while setting $\boldsymbol h_1=\boldsymbol h_2^*$ for $\boldsymbol h_2^*$ such that $-(\boldsymbol h_2^*) ^T \boldsymbol \theta_N^{\mathcal{H}}  + a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_2^*) = \min_{\boldsymbol h_2} [-\boldsymbol h_2^T \boldsymbol \theta_N^{\mathcal{H}}  + a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_2)]$ gives $\elt \geq 2 a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_2^*) \geq \Gamc \equiv \min_{\boldsymbol h}  a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h)$. Finally, note that for any $\boldsymbol h$, the triangle inequality gives
\begin{align*}
\Gam  \equiv \max_{\boldsymbol h_1}  a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h_1) &\geq
[a_{\boldsymbol \theta_N, \mathcal{H}} (\boldsymbol h)  + a_{\boldsymbol \theta_N, \mathcal{H}} (-\boldsymbol h)]/2 \\
&=
2^{-1}\sum_{i=1}^{N } \left(\left| \theta_{i,N}^{\mathcal{V}}   + \sum_{j=1}^{N_{\mathcal{H}}} h_j \theta_{ij,N}^{\mathcal{VH}} \right|
+ \left| \theta_{i,N}^{\mathcal{V}}   - \sum_{j=1}^{N_{\mathcal{H}}} h_j \theta_{ij,N}^{\mathcal{VH}} \right|\right)\\
& \geq \sum_{i=1}^{N }  \left| \theta_{i,N}^{\mathcal{V}} \right| \equiv |\boldsymbol \theta_{N}^{\mathcal{V}}|_1.
\end{align*}
\hfill $\Box$

\clearpage

# References {-}