### 2017-11-30 writing/paper-rbm/letters/cover_letter.Rmd

---
author: 
 - Andee Kaplan
 - Postdoctoral Associate
 - Department of Statistical Sciences 
 - Duke University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: Dear Drs. Liu and Zhao,
closing: Best regards,
address: 
 - Regina Liu and Hongyu Zhao
 - Co-Editors-Elect
 - "*Journal of the American Statistical Association - Theory & Methods*"
return-address:
 - Andee Kaplan
 - Department of Statistical Sciences
 - Duke University
 - P.O. Box 90251
 - Durham, NC 27708-0251
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

I am writing to submit an article entitled "Properties and Bayesian fitting of restricted Boltzmann machines" for consideration for publication in the Journal of the American Statistical Association - Theory & Methods.  In this submission, we discuss the relationship between parameter specification and unattractive model properties---degeneracy, instability and uninterpretability---in restricted Boltzmann machines and describe related difficulties that can arise in likelihood-based and Bayes fitting of such models.

This paper references results from another manuscript, an invited revision of which is presently under review by Technometrics. We have provided a blinded copy of this second manuscript as a supplement for referees.

Thank you for your consideration.